The Lipid Interactome

### 2025-02-12 Background/DataAnalysisAndStatistics.qmd
updated data analysis background page

* Studies included in this repository which used identical FRAG Pipe TMT pipeline:
	- Thomas et al., 2025[@thomasTrifunctionalLipidDerivatives2025]
	- Farley et al., 2024 (A)[@farleyTrifunctionalSphinganineNew2024a]
	- Farley et al., 2024 (B)[@farleyTrifunctionalFattyAcid2024]

@@

>>>>

Dr. Frank Stein developed and implemented the data analysis pipeline at the European Molecular Biosciences Laboratories in Heidelberg, Germany. The mass spectrometry (MS) data were processed using IsobarQuant[@mateusThermalProteomeProfiling2016] and Mascot V.24 (Matrix Science), following the methodology previously described by Farley et al. (2024, A), Farley et al. (2024, B), and Thomas et al (2025)[@farleyTrifunctionalSphinganineNew2024a; @farleyTrifunctionalFattyAcid2024; @thomasTrifunctionalLipidDerivatives2025]. To ensure robust protein identification, the analysis utilized a reverse UniProt FASTA database that included common contaminants. Proteins were only included in the final analysis if they contained at least two unique peptides, each with a minimum length of seven amino acids, and met a stringent false discovery rate (FDR) threshold of 0.01.

The acquired MS spectra were searched for specific peptide modifications, including fixed modifications (carbamidomethylation of cysteine and TMT-16-plex labeling of lysine) and variable modifications (N-terminal acetylation, methionine oxidation, and N-terminal TMT-16-plex labeling). Mass error tolerances were set to 10 ppm for full-scan MS spectra and 0.02 Da for MS/MS spectra, with a maximum of two missed cleavage sites permitted. To correct for batch effects across experimental replicates, a variance stabilization normalization was applied to log~2~-transformed raw data[@huberVarianceStabilizationApplied2002]. Changes in protein abundance between experimental and control conditions were assessed using LIMMA analysis[@ritchieLimmaPowersDifferential2015].

For the protein-lipid probe interaction experiments presented here (+/- UV exposure), proteins were classified as "enriched hits” if they showed at least a log~2~ fold-change of 1 toward the +UV condition with a p-value < 0.05. Proteins enriched under +UV conditions but with p-values > 0.05 were labeled as “enriched candidates.”

@@

* Need these details.

### 2025-02-12 BioinformaticsApplicationNote/GG_2025_applicationNote.qmd
updated data analysis background page

| Gene name | Enrichment value (logFC) | p value |  AveExpr   |  FDR   |
|:----------|:------------------------:|:-------:|:----------:|:------:|
| GeneA     |           1.2            |  0.001  | 1.0x10^18^ | 0.0001 |
| GeneB     |           0.1            |   0.6   | 1.0x10^5^  | 0.005  |
| ...       |           ...            |   ...   |    ...     |  ...   |
: **Example data table for smooth incorporation into Lipid Interactome.** ***Column descriptions:*** *Gene name*: The identifier for genes identified in the dataset (or Accession Number). *Enrichment value*: Fold change of Experimental versus Control fold-change, typically log2-transformed -- ideally the product of TMT or other quantitative mass spectrometry methods. *p value*: the results of an ANOVA or LIMMA analysis, corrected for multiple hypothesis testing. *AveExpr*: An averaged value of the m/z ion intensity of the Experimental condition and Control condition used for the enrichment value (used to prepare MA plots for quality control). *FDR*: False-discovery rate of protein identification. {#tbl-1 .striped}

>>>>

| Gene name | logFC | p value |  AveExpr   |  FDR   |
|:----------|:------------------------:|:-------:|:----------:|:------:|
| GeneA     |           1.2            |  0.001  | 1.0x10^18^ | 0.0001 |
| GeneB     |           0.1            |   0.6   | 1.0x10^5^  | 0.005  |
| ...       |           ...            |   ...   |    ...     |  ...   |
: **Example data table for smooth incorporation into Lipid Interactome.** ***Column descriptions:*** *Gene name*: The identifier for genes identified in the dataset (or Accession Number). *logFC*: Fold change of Experimental versus Control fold-change, typically log2-transformed -- ideally the product of TMT or other quantitative mass spectrometry methods. *p value*: the results of an ANOVA or LIMMA analysis, corrected for multiple hypothesis testing. *AveExpr*: An averaged value of the m/z ion intensity of the Experimental condition and Control condition used for the enrichment value (used to prepare MA plots for quality control). *FDR*: False-discovery rate of protein identification. {#tbl-1 .striped}
